#Chapter 2. Small Worlds and Large Worlds

```{r}
library(tidyverse)
```


## Notes


### 2.1.1. Counting Possiblities

```{r}
d <-
  tibble(p_1 = 0,
         p_2 = rep(1:0, times = c(1, 3)),
         p_3 = rep(1:0, times = c(2, 2)),
         p_4 = rep(1:0, times = c(3, 1)),
         p_5 = 1)

head(d)
```

```{r}
d %>% 
  pivot_longer(cols = 1:5) %>% 
  arrange(name) %>% 
  mutate(x = rep(1:4, times = 5),
         possibility = rep(1:5, each = 4)) %>% 
  
  ggplot(aes(x = x, y = possibility, 
             fill = value %>% as.character())) +
  geom_point(shape = 21, size = 5) +
  scale_fill_manual(values = c("white", "navy")) +
  scale_x_continuous(NULL, breaks = NULL) +
  coord_cartesian(xlim = c(.75, 4.25),
                  ylim = c(.75, 5.25)) +
  theme(legend.position = "none")
```


```{r}
tibble(draw            = 1:3,
       marbles         = 4) %>% 
  mutate(possibilities = marbles ^ draw) %>% 
  knitr::kable()
```

```{r}
(
  d <-
  tibble(position = c((1:4^1) / 4^0, 
                      (1:4^2) / 4^1, 
                      (1:4^3) / 4^2),
         draw     = rep(1:3, times = c(4^1, 4^2, 4^3)),
         fill     = rep(c("b", "w"), times = c(1, 3)) %>% 
           rep(., times = c(4^0 + 4^1 + 4^2)))
  )
```


```{r}
d %>% 
  ggplot(aes(x = position, y = draw)) +
  geom_point(aes(fill = fill),
             shape = 21, size = 3) +
  scale_y_continuous(breaks = 1:3) +
  scale_fill_manual(values  = c("navy", "white")) +
  theme(panel.grid.minor = element_blank(),
        legend.position  = "none")

(
  lines_1 <-
  tibble(x    = rep((1:4), each = 4),
         xend = ((1:4^2) / 4),
         y    = 1,
         yend = 2)
  )

(
  lines_2 <-
  tibble(x    = rep(((1:4^2) / 4), each = 4),
         xend = (1:4^3) / (4^2),
         y    = 2,
         yend = 3)
  )

d %>% 
  ggplot(aes(x = position, y = draw)) +
  geom_segment(data  = lines_1,
               aes(x = x, xend = xend,
                   y = y, yend = yend),
               size  = 1/3) +
  geom_segment(data  = lines_2,
               aes(x = x, xend = xend,
                   y = y, yend = yend),
               size  = 1/3) +
  geom_point(aes(fill = fill),
             shape = 21, size = 3) +
  scale_y_continuous(breaks = 1:3) +
  scale_fill_manual(values  = c("navy", "white")) +
  theme(panel.grid.minor = element_blank(),
        legend.position  = "none")

d <-
  d %>% 
  mutate(denominator = ifelse(draw == 1, .5,
                              ifelse(draw == 2, .5 / 4,
                                     .5 / 4^2))) %>% 
  mutate(position    = position - denominator)

d

(
  lines_1 <-
  lines_1 %>% 
  mutate(x    = x - .5,
         xend = xend - .5 / 4^1)
)

(
  lines_2 <-
  lines_2 %>% 
  mutate(x    = x - .5 / 4^1,
         xend = xend - .5 / 4^2)
)

d %>% 
  ggplot(aes(x = position, y = draw)) +
  geom_segment(data  = lines_1,
               aes(x = x, xend = xend,
                   y = y, yend = yend),
               size  = 1/3) +
  geom_segment(data  = lines_2,
               aes(x = x, xend = xend,
                   y = y, yend = yend),
               size  = 1/3) +
  geom_point(aes(fill = fill),
             shape = 21, size = 3) +
  scale_y_continuous(breaks = 1:3) +
  scale_fill_manual(values  = c("navy", "white")) +
  theme(panel.grid.minor = element_blank(),
        legend.position  = "none")


d %>% 
  ggplot(aes(x = position, y = draw)) +
  geom_segment(data  = lines_1,
               aes(x = x, xend = xend,
                   y = y, yend = yend),
               size  = 1/3) +
  geom_segment(data  = lines_2,
               aes(x = x, xend = xend,
                   y = y, yend = yend),
               size  = 1/3) +
  geom_point(aes(fill = fill),
             shape = 21, size = 4) +
  scale_fill_manual(values  = c("navy", "white")) +
  scale_x_continuous(NULL, limits = c(0, 4), breaks = NULL) +
  scale_y_continuous(NULL, limits = c(0.75, 3), breaks = NULL) +
  theme(panel.grid      = element_blank(),
        legend.position = "none") +
  coord_polar()

lines_1 <-
  lines_1 %>% 
  mutate(remain = c(rep(0:1, times = c(1, 3)),
                    rep(0,   times = 4 * 3)))

lines_2 <-
  lines_2 %>% 
  mutate(remain = c(rep(0,   times = 4),
                    rep(1:0, times = c(1, 3)) %>% 
                      rep(., times = 3),
                    rep(0,   times = 12 * 4)))

d <-
  d %>% 
  mutate(remain = c(rep(1:0, times = c(1, 3)),
                    rep(0:1, times = c(1, 3)),
                    rep(0,   times = 4 * 4),
                    rep(1:0, times = c(1, 3)) %>% 
                      rep(., times = 3),
                    rep(0,   times = 12 * 4))) 

# finally, the plot:
d %>% 
  ggplot(aes(x = position, y = draw)) +
  geom_segment(data  = lines_1,
               aes(x = x, xend = xend,
                   y = y, yend = yend,
                   alpha = remain %>% as.character()),
               size  = 1/3) +
  geom_segment(data  = lines_2,
               aes(x = x, xend = xend,
                   y = y, yend = yend,
                   alpha = remain %>% as.character()),
               size  = 1/3) +
  geom_point(aes(fill = fill, alpha = remain %>% as.character()),
             shape = 21, size = 4) +
  # it's the alpha parameter that makes elements semitransparent
  scale_alpha_manual(values = c(1/10, 1)) +
  scale_fill_manual(values  = c("navy", "white")) +
  scale_x_continuous(NULL, limits = c(0, 4), breaks = NULL) +
  scale_y_continuous(NULL, limits = c(0.75, 3), breaks = NULL) +
  theme(panel.grid      = element_blank(),
        legend.position = "none") +
  coord_polar()


n_blue <- function(x){
  rowSums(x == "b")
}

n_white <- function(x){
  rowSums(x == "w")
}

t <-
  # for the first four columns, `p_` indexes position
  tibble(p_1 = rep(c("w", "b"), times = c(1, 4)),
         p_2 = rep(c("w", "b"), times = c(2, 3)),
         p_3 = rep(c("w", "b"), times = c(3, 2)),
         p_4 = rep(c("w", "b"), times = c(4, 1))) %>% 
  mutate(`draw 1: blue`  = n_blue(.),
         `draw 2: white` = n_white(.),
         `draw 3: blue`  = n_blue(.)) %>% 
  mutate(`ways to produce` = `draw 1: blue` * `draw 2: white` * `draw 3: blue`)

t %>% 
  knitr::kable()
```

```{r}
d <-
  tibble(position = c((1:4^1) / 4^0, 
                      (1:4^2) / 4^1, 
                      (1:4^3) / 4^2),
         draw     = rep(1:3, times = c(4^1, 4^2, 4^3)))

(
  d <-
  d %>% 
  bind_rows(
    d, d
  ) %>% 
  # here are the fill colors
  mutate(fill = c(rep(c("w", "b"), times = c(1, 3)) %>% rep(., times = c(4^0 + 4^1 + 4^2)),
                  rep(c("w", "b"), each  = 2)       %>% rep(., times = c(4^0 + 4^1 + 4^2)),
                  rep(c("w", "b"), times = c(3, 1)) %>% rep(., times = c(4^0 + 4^1 + 4^2)))) %>% 
  # now we need to shift the positions over in accordance with draw, like before
  mutate(denominator = ifelse(draw == 1, .5,
                              ifelse(draw == 2, .5 / 4,
                                     .5 / 4^2))) %>% 
  mutate(position = position - denominator) %>% 
  # here we'll add an index for which pie wedge we're working with
  mutate(pie_index = rep(letters[1:3], each = n()/3)) %>% 
  # to get the position axis correct for pie_index == "b" or "c", we'll need to offset
  mutate(position = ifelse(pie_index == "a", position,
                           ifelse(pie_index == "b", position + 4,
                                  position + 4 * 2)))
  )
```

```{r}
move_over <- function(position, index){
  ifelse(index == "a", position,
         ifelse(index == "b", position + 4,
                position + 4 * 2)
         )
}

(
  lines_1 <-
  tibble(x    = rep((1:4), each = 4) %>% rep(., times = 3),
         xend = ((1:4^2) / 4)        %>% rep(., times = 3),
         y    = 1,
         yend = 2) %>% 
  mutate(x    = x - .5,
         xend = xend - .5 / 4^1) %>% 
  # here we'll add an index for which pie wedge we're working with
  mutate(pie_index = rep(letters[1:3], each = n()/3)) %>% 
  # to get the position axis correct for `pie_index == "b"` or `"c"`, we'll need to offset
  mutate(x    = move_over(position = x,    index = pie_index),
         xend = move_over(position = xend, index = pie_index))
  )

(
  lines_2 <-
  tibble(x    = rep(((1:4^2) / 4), each = 4)  %>% rep(., times = 3),
         xend = (1:4^3 / 4^2)                 %>% rep(., times = 3),
         y    = 2,
         yend = 3) %>% 
  mutate(x    = x - .5 / 4^1,
         xend = xend - .5 / 4^2) %>% 
  # here we'll add an index for which pie wedge we're working with
  mutate(pie_index = rep(letters[1:3], each = n()/3)) %>% 
  # to get the position axis correct for `pie_index == "b"` or `"c"`, we'll need to offset
  mutate(x    = move_over(position = x,    index = pie_index),
         xend = move_over(position = xend, index = pie_index))
  )

d <- 
  d %>% 
  mutate(remain = c(# `pie_index == "a"`
                    rep(0:1, times = c(1, 3)),
                    rep(0,   times = 4),
                    rep(1:0, times = c(1, 3)) %>% 
                      rep(., times = 3),
                    rep(0,   times = 4 * 4),
                    rep(c(0, 1, 0), times = c(1, 3, 4 * 3)) %>% 
                      rep(., times = 3),
                    # `pie_index == "b"`
                    rep(0:1, each = 2),
                    rep(0,   times = 4 * 2),
                    rep(1:0, each = 2) %>% 
                      rep(., times = 2),
                    rep(0,   times = 4 * 4 * 2),
                    rep(c(0, 1, 0, 1, 0), times = c(2, 2, 2, 2, 8)) %>% 
                      rep(., times = 2),
                    # `pie_index == "c"`
                    rep(0:1, times = c(3, 1)),
                    rep(0,   times = 4 * 3),
                    rep(1:0, times = c(3, 1)), 
                    rep(0,   times = 4 * 4 * 3),
                    rep(0:1, times = c(3, 1)) %>% 
                      rep(., times = 3),
                    rep(0,   times = 4)
                    )
         )

lines_1 <-
  lines_1 %>% 
  mutate(remain = c(rep(0,   times = 4),
                    rep(1:0, times = c(1, 3)) %>% 
                      rep(., times = 3),
                    rep(0,   times = 4 * 2),
                    rep(1:0, each  = 2) %>% 
                      rep(., times = 2),
                    rep(0,   times = 4 * 3),
                    rep(1:0, times = c(3, 1))
                    )
         )

lines_2 <-
  lines_2 %>% 
  mutate(remain = c(rep(0,   times = 4 * 4),
                    rep(c(0, 1, 0), times = c(1, 3, 4 * 3)) %>% 
                      rep(., times = 3),
                    rep(0,   times = 4 * 8),
                    rep(c(0, 1, 0, 1, 0), times = c(2, 2, 2, 2, 8)) %>% 
                      rep(., times = 2),
                    rep(0,   times = 4 * 4 * 3),
                    rep(0:1, times = c(3, 1)) %>% 
                      rep(., times = 3),
                    rep(0,   times = 4)
                    )
         )
```


```{r}
d %>% 
  ggplot(aes(x = position, y = draw)) +
  geom_vline(xintercept = c(0, 4, 8), color = "white", size = 2/3) +
  geom_segment(data  = lines_1,
               aes(x = x, xend = xend,
                   y = y, yend = yend,
                   alpha = remain %>% as.character()),
               size  = 1/3) +
  geom_segment(data  = lines_2,
               aes(x = x, xend = xend,
                   y = y, yend = yend,
                   alpha = remain %>% as.character()),
               size  = 1/3) +
  geom_point(aes(fill = fill, size = draw, alpha = remain %>% as.character()),
             shape = 21) +
  scale_size_continuous(range = c(3, 1.5)) +
  scale_alpha_manual(values = c(1/10, 1)) +
  scale_fill_manual(values  = c("navy", "white")) +
  scale_x_continuous(NULL, limits = c(0, 12),     breaks = NULL) +
  scale_y_continuous(NULL, limits = c(0.75, 3.5), breaks = NULL) +
  theme(panel.grid      = element_blank(),
        legend.position = "none") +
  coord_polar()
```



```{r}
ways <- c(0, 3, 8, 9, 0)
ways/sum(ways)
```


```{r}
dbinom( 6 , size=9 , prob=0.5 )
```

```{r}
# define grid
p_grid <- seq( from=0 , to=1 , length.out=20 )
# define prior
prior <- rep( 1 , 20 )
# compute likelihood at each value in grid
likelihood <- dbinom( 6 , size=9 , prob=p_grid )
# compute product of likelihood and prior
unstd.posterior <- likelihood * prior
# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)
```

```{r}
plot( p_grid , posterior , type="b" ,
    xlab="probability of water" , ylab="posterior probability" )
mtext( "20 points" )
```

```{r}
prior <- ifelse( p_grid < 0.5 , 0 , 1 )
likelihood <- dbinom( 6 , size=9 , prob=p_grid )
# compute product of likelihood and prior
unstd.posterior <- likelihood * prior
# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)
plot( p_grid , posterior , type="b" ,
    xlab="probability of water" , ylab="posterior probability" )
mtext( "20 points" )


prior <- exp( -5*abs( p_grid - 0.5 ) )
likelihood <- dbinom( 6 , size=9 , prob=p_grid )
# compute product of likelihood and prior
unstd.posterior <- likelihood * prior
# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)
plot( p_grid , posterior , type="b" ,
    xlab="probability of water" , ylab="posterior probability" )
mtext( "20 points" )
```



## Excercises


2E1. Which of the expressions below correspond to the statement: the probability of rain on Monday?

- (1) Pr(rain)
- **(2) Pr(rain|Monday)**
- (3) Pr(Monday|rain)
- **(4) Pr(rain,Monday)/Pr(Monday)**

Given that it is monday, what is the probaility that it rains.

$$
Pr(Rain|Monday) = \frac{Pr(rain, Monday)}{Pr(Monday)}
$$

:::question
> **2E2.** Which of the following statements corresponds to the expression: Pr(Monday|rain)?  
>   (1) The probability of rain on Monday.  
>   (2) The probability of rain, given that it is Monday.  
>   **(3) The probability that it is Monday, given that it is raining.  **
>   (4) The probability that it is Monday and that it is raining.
:::


