#Chapter 2. Small Worlds and Large Worlds

## Notes

```{r}
ways <- c(0, 3, 8, 9, 0)
ways/sum(ways)
```


```{r}
dbinom( 6 , size=9 , prob=0.5 )
```

```{r}
# define grid
p_grid <- seq( from=0 , to=1 , length.out=20 )
# define prior
prior <- rep( 1 , 20 )
# compute likelihood at each value in grid
likelihood <- dbinom( 6 , size=9 , prob=p_grid )
# compute product of likelihood and prior
unstd.posterior <- likelihood * prior
# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)
```

```{r}
plot( p_grid , posterior , type="b" ,
    xlab="probability of water" , ylab="posterior probability" )
mtext( "20 points" )
```

```{r}
prior <- ifelse( p_grid < 0.5 , 0 , 1 )
likelihood <- dbinom( 6 , size=9 , prob=p_grid )
# compute product of likelihood and prior
unstd.posterior <- likelihood * prior
# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)
plot( p_grid , posterior , type="b" ,
    xlab="probability of water" , ylab="posterior probability" )
mtext( "20 points" )


prior <- exp( -5*abs( p_grid - 0.5 ) )
likelihood <- dbinom( 6 , size=9 , prob=p_grid )
# compute product of likelihood and prior
unstd.posterior <- likelihood * prior
# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)
plot( p_grid , posterior , type="b" ,
    xlab="probability of water" , ylab="posterior probability" )
mtext( "20 points" )
```



## Excercises


2E1. Which of the expressions below correspond to the statement: the probability of rain on Monday?

- (1) Pr(rain)
- **(2) Pr(rain|Monday)**
- (3) Pr(Monday|rain)
- **(4) Pr(rain,Monday)/Pr(Monday)**

Given that it is monday, what is the probaility that it rains.

$$
Pr(Rain|Monday) = \frac{Pr(rain, Monday)}{Pr(Monday)}
$$

:::question
> **2E2.** Which of the following statements corresponds to the expression: Pr(Monday|rain)?  
>   (1) The probability of rain on Monday.  
>   (2) The probability of rain, given that it is Monday.  
>   **(3) The probability that it is Monday, given that it is raining.  **
>   (4) The probability that it is Monday and that it is raining.
:::


